\section{Motivation and Design Objectives}
\label{sec:heuristics_motivation}
\hspace{\parindent}
The motivation for developing prioritization heuristics arises from the fact that the same planning instance may produce substantially different results depending only on the order in which robots are planned. In constrained environments, assigning priority to one robot may reduce its own planning difficulty while increasing the constraints imposed on the remaining robots.

Since the most suitable ordering depends on the spatial distribution of robots, their goals, the structure of the roadmap, and the interactions between paths, a single fixed prioritization rule is unlikely to perform consistently across all scenarios. For this reason, the objective of this work is not to identify one dominant heuristic, but to design a diverse set of complementary ordering strategies. These heuristics should produce priority orderings that are informative enough to improve planning feasibility or solution quality, while remaining computationally lightweight when compared with the execution time of TEA*. This cost requirement is particularly important in the context of online fleet coordination. Since prioritization is performed before planner execution, excessive heuristic computation time could reduce or eliminate the practical benefit obtained from improved orderings.

\section{Heuristic Design Methodology}
\label{sec:heuristic_design_methodology}
\hspace{\parindent}
The heuristics developed in this work follow a common score-based formulation. Given a set of robots

\begin{equation}
    \mathcal{R} = \{r_1, r_2, \ldots, r_n\}
\end{equation}
, each heuristic assigns a priority score, $score_i$, to every robot, $r_i$, according to the information available in the current planning instance. For a given heuristic, this score is denoted as:

\begin{equation}
    score_i = h(r_i, \mathcal{P})
\end{equation}
, where $\mathcal{P}$ represents the current planning instance, including information such as the robots' current positions, target vertices, roadmap graph properties, estimated paths, predicted conflicts, or information from previous coordination cycles.

The final priority ordering is obtained by sorting the robots according to their scores. Depending on the heuristic, robots with higher scores may be planned first or last.
This formulation provides a common implementation interface for all proposed heuristics. Each heuristic differs in the information used to compute $score_i$ and in the direction of the sorting operation. As a result, different prioritization strategies can be used within the same TEA* planning pipeline and later explored in parallel.

The developed heuristics are organized into five main groups according to the type of information they exploit:

\begin{itemize}
    \item \textbf{Roadmap distance heuristics}, based on distances or estimated path lengths in the roadmap graph;

    \item \textbf{Topological heuristics}, based on structural properties of the graph and its constrained regions;

    \item \textbf{Surroundings-based heuristics}, based on the spatial distribution of robots around their current positions or target goals;

    \item \textbf{Conflict-based heuristics}, based on estimated interactions between tentative robot paths;

    \item \textbf{Replanning heuristics}, based on information from previous plans or coordination cycles.
\end{itemize}

Together, these groups cover complementary dimensions of the coordination problem, allowing the proposed heuristics to address robot prioritization from different perspectives. Each group is examined in the following sections, where the corresponding heuristics are described in terms of their motivation, scoring procedure, and computational implications.

\section{Roadmap Distance Heuristics}
\label{sec:roadmap_distance_heuristics}
\hspace{\parindent}
Roadmap distance heuristics define robot priority orderings using distance-related measures between each robot and its target. The main intuition behind this family is that robots associated with longer movements may have a higher probability of interacting with other robots, since their paths tend to cover a larger portion of the environment and may traverse more shared resources.

These heuristics have the advantage of being computationally lightweight, as the priority permutation is computed from simple distance measures or shortest-path estimates. However, they do not explicitly account for path conflicts, local robot density, or topological constraints. Instead, they use distance to the target as a simple proxy for planning difficulty.

In this work, two distance-based criteria are considered, namely the distance between each robot and its target and the length of an estimated path in the roadmap graph. These criteria give rise to the Furthest Robot and Longest Path heuristics, respectively.


\subsection{Furthest Robot}
\label{subsec:furthest_robot}

\subsection{Longest Path}
\label{subsec:longest_path}


\newpage
\section{Topological Heuristics}
\label{sec:topological_heuristics}

\section{Surroundings-based Heuristics}
\label{sec:surroundings_heuristics}

\section{Conflicts-based Heuristics}
\label{sec:conflicts_heuristics}

\section{Replanning Heuristics}
\label{sec:replanning_heuristics}
