\section{MRPP Problem and MAPF Abstraction}
\label{sec:problem-definition}
\hspace{\parindent}
The problem addressed in this dissertation is centralized MRPP, where
multiple robots must move from their initial states to their target states while sharing the same
environment. Each robot must reach its goal without colliding with other robots, and the resulting
set of trajectories should satisfy a global performance criterion.

A widely used abstraction for formalizing this type of coordination problem is MAPF. In MAPF, the moving entities are usually referred to as agents, but in the context
of this dissertation they correspond to robots in a fleet coordination system. The formulation
adopted here follows the conceptual definition of MAPF presented by Stern et
al.~\cite{stern2019definitions}, while being interpreted in the MRPP setting considered throughout
this work.

From a formal perspective, let there be a set of \(m\) robots \(R = \{r_1, r_2, \dots, r_m\}\).
Each robot \(r_i \in R\) is associated with an initial state \(s_i\) and a goal state \(g_i\).

A solution for robot \(r_i\) is defined as a trajectory \(P_i\), which specifies the evolution of the
robot state over time. In discrete formulations, a trajectory can be represented as a sequence of
states
\begin{equation}
    P_i = \langle p_i^0, p_i^1, \dots, p_i^T \rangle
\end{equation}
where \(p_i^0 = s_i\) and \(p_i^T = g_i\). The explicit consideration of time is essential, since
interactions and conflicts between robots arise from the simultaneous execution of their
trajectories in a shared space-time domain~\cite{stern2019definitions}.

A valid solution must ensure that no two robots are in conflict during their motion, where conflicts
arise when robots attempt to occupy incompatible states in space and time. The most common type of
conflict is the position conflict, which occurs when two robots occupy the same state at the same
time step, i.e., \(p_i^t = p_j^t\) for \(i \neq j\). Other types of conflicts may arise depending on
the motion model and representation adopted, but all correspond to violations of mutual feasibility
constraints among robots~\cite{stern2019definitions}.

The objective is to compute a set of conflict-free trajectories
\(\{P_1, P_2, \dots, P_m\}\) that minimize a global cost function \(C\). This cost function is
defined over the joint set of robot trajectories and encodes the chosen optimization criterion.
Common objectives include minimizing the time at which the last robot reaches its goal, known as
the makespan, or minimizing the sum of individual traversal costs. Formally:

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!30, boxrule=0.3pt, arc=3pt, left=3pt, right=3pt, top=1pt, bottom=1pt]
\vspace{-3pt}
\[
\textbf{Find } \{P_1, P_2, \dots, P_m\} \textbf{ such that all trajectories are conflict-free and } C(\{P_i\}) \textbf{ is minimized.}
\]
\vspace{-6pt}
\end{tcolorbox}

This formulation inherits the combinatorial nature of MAPF. In the general case, optimal MAPF is
NP-hard, as formally established in earlier works~\cite{surynek2010,yu2013planning} and confirmed
in recent complexity analyses~\cite{geft2022refined,gao2024review}. This inherent complexity
motivates the development of planning algorithms that balance solution quality with computational
efficiency in practical MRPP systems.