\chapter{Temporal Consistency and Redundant Replanning Mitigation through Batching} 
\label{chap:temporal_consistency_batching}
\label{chap:ch9}
\hspace{\parindent}
The previous chapters addressed the performance of the Path Planner by studying priority
heuristics and by introducing a parallel planning framework capable of evaluating multiple
heuristic-guided orderings within the available planning time. These contributions were evaluated
using an isolated Path Planner testing framework, which made it possible to analyze the planning
process independently from the rest of the coordination stack.

However, improving the planner alone does not fully determine the behavior of the complete CRIIS
coordination system during integrated execution. When the full stack is executed, the Path Planner
operates as part of a broader feedback loop involving the Navigation Handlers and the Supervisor.
In this context, planning requests, robot state updates, and execution feedback may be generated
within short time intervals, creating situations that are not captured by isolated planner tests.

This chapter therefore shifts the analysis from planning performance to coordination consistency.
During integrated executions, temporal consistency issues were observed in the processing of
planning requests and robot state updates, which can lead to redundant replanning operations.
To mitigate these effects, batching mechanisms were introduced to group closely related planning
requests and robot state updates before they are processed by the coordination pipeline.

\input{chapters/9.baseline_system_stabilization/sections/initial_issues.tex}
\input{chapters/9.baseline_system_stabilization/sections/temporal_analysis.tex}