\section{Document Structure}
\label{sec:document_structure}
\hspace{\parindent}
This dissertation is organized as follows. Chapter~\ref{chap:ch2} presents the fundamental
concepts required to support the work, including MRPP, the MAPF abstraction, graph search,
machine learning concepts for clustering and runtime analysis, and parallel computing.
Chapter~\ref{chap:ch3} reviews the state of the art in multi-robot coordination and related MAPF
approaches.

Chapter~\ref{chap:ch4} describes the CRIIS fleet coordination system used as the starting point for
this work, including its centralized planning pipeline and the role of TEA* as the underlying path
planning method.

Chapter~\ref{chap:ch5} presents the proposed prioritization heuristics, which generate alternative
robot orderings for TEA* from different sources of information, including distance, roadmap
structure, local robot distribution, predicted conflicts, and replanning feedback. Their individual
behaviour is evaluated in Chapter~\ref{chap:ch6} using an isolated Path Planner testing framework,
allowing the effect of each ordering strategy to be analysed independently from the remaining
coordination stack.

Chapter~\ref{chap:ch7} introduces the proposed parallel planning framework, including the thread
worker implementation, heuristic dispatching mechanism, timeout handling, and redundant heuristic
selection strategy. The performance of this framework is then assessed in Chapter~\ref{chap:ch8},
with emphasis on the effect of the number of worker threads and on the contribution of the
heuristic dispatching strategy.

Finally, Chapter~\ref{chap:ch9} analyses temporal consistency issues observed during execution of
the complete CRIIS coordination stack, focusing on redundant replanning operations and on the
batching mechanisms introduced for planning requests and robot state updates.
Chapter~\ref{chap:ch10} summarizes the main conclusions and outlines possible directions for
future work.
