\section{Contributions}
\hspace{\parindent}
This dissertation builds upon the centralized fleet coordination system described by
Matos et al.~\cite{matos2025criis} and proposes an extension of the sequential TEA* planner aimed
at reducing its sensitivity to robot priority ordering. One contribution is the development of a set
of prioritization heuristics that generate alternative robot orderings from different properties of the
planning instance, providing structured alternatives to random priority selection.

A second contribution is the integration of these heuristics into a multithreaded planning approach,
where multiple TEA* instances explore different priority orderings within the same planning time.
The goal is to assess whether more robust and lower-cost coordination solutions can be obtained without modifying the
core architecture of the existing system.

By evaluating a broader and more informative set of priority orderings, the proposed approach aims
to improve performance in the constrained multi-robot scenarios considered in this dissertation, where the ordering strongly affects
solution feasibility. Additionally, this work provides empirical insight into the role of heuristic
prioritization and parallel execution in sequential planning, contributing to the design of future
centralized fleet coordination systems.

An additional outcome of this dissertation is the preparation and submission of a scientific paper to
the 9th Iberian Robotics Conference, ROBOT2026. The submitted work reports part of the heuristic
prioritization study developed in this dissertation, contributing to the dissemination of the proposed
approach within the robotics research community.