\chapter*{Acknowledgements}
\hspace{\parindent}
I would like to express my sincere gratitude to Professor Pedro Costa, my supervisor, for his guidance, availability, and constant support throughout this dissertation. His advice helped me follow the right direction at important moments, and his calm and practical approach made the process clearer and less overwhelming.

I am also deeply grateful to Diogo Matos, my second supervisor and advisor at INESC TEC, for his continuous presence and support. His guidance, technical insight, and readiness to discuss ideas and decisions were essential throughout the development of this dissertation.

To Marco Moreira, my colleague in the mobile robotics team, I would like to express a special thank you for his friendship, good humor, and constant willingness to help throughout this journey. His presence made this process much more enjoyable.

I would also like to thank Héber Sobreira for always being available, supportive, and ready to offer good advice. His good humor and reassuring presence made many moments of this journey lighter.


To all members of the mobile robotics group, thank you for the companionship, collaboration, and mutual support during these months. Being surrounded by people who were always willing to help and share ideas made this experience much more meaningful.

I also want to thank all members of the 5DPO robotics team, with whom I shared a journey that I will always remember. More than teammates, many became friends for life. In particular, I would like to thank Daniel Silva and Pedro Lopes for all the work, challenges, laughter, and moments we shared together.
I am also grateful to Professor António Paulo Moreira and Professor Paulo Costa for the support they always showed, both in curricular matters and in the extracurricular path of the team.

To Leonor, my love, I cannot thank you enough. Your patience and constant support carried me through the most difficult moments. Thank you for believing in me when I doubted myself, for listening, and for always being by my side. Without you, this path would have been much harder.

To my father José, my mother Rosa and my sister Diana, thank you for being my foundation. Your support and trust have always given me the strength to move forward. Everything I achieve is also a reflection of what you have given me, and this dissertation would not have been possible without you.

Finally, to my friends, thank you for being present and for helping me disconnect when I needed it most. Your support and friendship were essential to keep balance throughout this period.

\vspace{10mm}
\flushleft{Afonso Mateus}
